---
title: "masters_thesis_air_pollution_schools_results"
author: "Steven Futter"
date: "7/15/2017"
output: html_document
---

_ann_test_8_hr_ahead_reset_13.csv

# TEST SET
# 1. LAGS 1-3, CURRENT METEO, SLOPES PRIOR 8AM
# MAE & RESULTS TABLE FOR TEST 
```{r}
infilePath = file.path("~/Dropbox","NU","THESIS","DATASETS","RESULTS", "PREDICTIONS","ONE_HOUR_AHEAD_PREDICTIONS")
df         = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_1_hr_ahead.csv"),na.strings=c("NA"," "))
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_2_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_3_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_4_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_5_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_6_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_7_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_8_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
df

# CALC the MAE here
# TESTING: evaluate the quality of the prediction using the test data set:
df.test.mae = mean(abs(df$predict-df$NO2))
df.test.mae          # 4.92347 NO2

# CONFUSION MATRIX
df$predictOver40 = ifelse(df$predict > 40, 'Predict Above 40', 'Predict Below 40')
df$actualOver40 = ifelse(df$NO2 > 40,  'Actual Above 40', 'Actual Below 40')
results = df %>% select(Time, predict, NO2, predictOver40, actualOver40)
table(results$predictOver40, results$actualOver40)


# accurately predict above 40 when actual is above 40   ***** GOOD NEWS IF THE MODEL IS GOOD AT THIS  ---- NEED TO MAXIMIZE *****
accurate_pred_over_40 = table(results$predictOver40, results$actualOver40)[1,1] / (table(results$predictOver40, results$actualOver40)[1,1] + table(results$predictOver40, results$actualOver40)[2,1])
paste0('Accurately predict NO2 above 40: ',round(accurate_pred_over_40,3))

# accurate predict below 40 when actual is below 40
accurate_pred_below_40 = table(results$predictOver40, results$actualOver40)[2,2] / (table(results$predictOver40, results$actualOver40)[1,2] + table(results$predictOver40, results$actualOver40)[2,2])
paste0('Accurately predict NO2 below 40: ',round(accurate_pred_below_40,3))

# inaccurately predict above 40 when actual is below 40
inaccurate_pred_above_40 = table(results$predictOver40, results$actualOver40)[1,2] / (table(results$predictOver40, results$actualOver40)[1,1] + table(results$predictOver40, results$actualOver40)[1,2])
paste0('Inaccurately predict NO2 above 40: ',round(inaccurate_pred_above_40,3))

# inaccurate predict below 40 when actual is above 40 ****BIG PROBLEM IS THIS IS THE CASE ---- NEED TO MINIMIZE THIS PERCENTAGE ******
inaccurate_pred_below_40 = table(results$predictOver40, results$actualOver40)[2,1] / (table(results$predictOver40, results$actualOver40)[2,1] + table(results$predictOver40, results$actualOver40)[1,1])
paste0('Inaccurately predict NO2 below 40: ',round(inaccurate_pred_below_40,3))

```


# VALI SET
# 1. LAGS 1-3, CURRENT METEO, SLOPES PRIOR 8AM
# MAE & RESULTS TABLE FOR VALIDATION
```{r}
# REPEAT FOR VALIDATION DATA SET
infilePath = file.path("~/Dropbox","NU","THESIS","DATASETS","RESULTS", "PREDICTIONS","ONE_HOUR_AHEAD_PREDICTIONS")
df         = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_1_hr_ahead.csv"),na.strings=c("NA"," "))
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_2_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_3_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_4_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_5_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_6_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_7_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_8_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
df

# CALC the MAE here
# VALIDATION: evaluate the quality of the prediction using the test data set
df.vali.mae = mean(abs(df$predict-df$NO2))
df.vali.mae          # 5.847267 NO2


# CONFUSION MATRIX
df$predictOver40 = ifelse(df$predict > 40, 'Predict Above 40', 'Predict Below 40')
df$actualOver40 = ifelse(df$NO2 > 40,  'Actual Above 40', 'Actual Below 40')
results = df %>% select(Time, predict, NO2, predictOver40, actualOver40)
table(results$predictOver40, results$actualOver40)


# accurately predict above 40 when actual is above 40   ***** GOOD NEWS IF THE MODEL IS GOOD AT THIS  ---- NEED TO MAXIMIZE *****
accurate_pred_over_40 = table(results$predictOver40, results$actualOver40)[1,1] / (table(results$predictOver40, results$actualOver40)[1,1] + table(results$predictOver40, results$actualOver40)[2,1])
paste0('Accurately predict NO2 above 40: ',round(accurate_pred_over_40,3))

# accurate predict below 40 when actual is below 40
accurate_pred_below_40 = table(results$predictOver40, results$actualOver40)[2,2] / (table(results$predictOver40, results$actualOver40)[1,2] + table(results$predictOver40, results$actualOver40)[2,2])
paste0('Accurately predict NO2 below 40: ',round(accurate_pred_below_40,3))

# inaccurately predict above 40 when actual is below 40
inaccurate_pred_above_40 = table(results$predictOver40, results$actualOver40)[1,2] / (table(results$predictOver40, results$actualOver40)[1,1] + table(results$predictOver40, results$actualOver40)[1,2])
paste0('Inaccurately predict NO2 above 40: ',round(inaccurate_pred_above_40,3))

# inaccurate predict below 40 when actual is above 40 ****BIG PROBLEM IS THIS IS THE CASE ---- NEED TO MINIMIZE THIS PERCENTAGE ******
inaccurate_pred_below_40 = table(results$predictOver40, results$actualOver40)[2,1] / (table(results$predictOver40, results$actualOver40)[2,1] + table(results$predictOver40, results$actualOver40)[1,1])
paste0('Inaccurately predict NO2 below 40: ',round(inaccurate_pred_below_40,3))

```


# DRAW PLOTLY OF DAILY PREDICT VS ACTUAL FOR DAYS WHEN OVER 40 ... 
# DRAW TABLE WITH NUMBER OF TIMES >40 WHERE PREDICT ACCURATE
```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[1])

good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
good.predict$text = 'Accurate Prediction'
bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[1])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```




```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[2])

good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
good.predict$text = 'Accurate Prediction'
bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[2])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```



```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[3])

good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
good.predict$text = 'Accurate Prediction'
bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[3])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```




```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[4])

good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
good.predict$text = 'Accurate Prediction'
#bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
#bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[4])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```



```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[5])

good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
good.predict$text = 'Accurate Prediction'
#bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
#bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[5])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```



```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[6])

#good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
#good.predict$text = 'Accurate Prediction'
bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[6])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```

















################################################################
########## RESET WITH 1PM PREDICTION USING ACTUAL DATA #########
################################################################

# TEST SET - (with reset at 1pm)
# 1. LAGS 1-3, CURRENT METEO, SLOPES PRIOR 8AM
# MAE & RESULTS TABLE FOR TEST 
```{r}
infilePath = file.path("~/Dropbox","NU","THESIS","DATASETS","RESULTS", "PREDICTIONS","ONE_HOUR_AHEAD_PREDICTIONS")
df         = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_1_hr_ahead.csv"),na.strings=c("NA"," "))
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_2_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_3_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_4_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_5_hr_ahead_reset_13.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_6_hr_ahead_reset_13.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_7_hr_ahead_reset_13.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_test_8_hr_ahead_reset_13.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
df

# CALC the MAE here
# TESTING: evaluate the quality of the prediction using the test data set:
df.test.mae = mean(abs(df$predict-df$NO2))
df.test.mae          # 4.92347 NO2

# CONFUSION MATRIX
df$predictOver40 = ifelse(df$predict > 40, 'Predict Above 40', 'Predict Below 40')
df$actualOver40 = ifelse(df$NO2 > 40,  'Actual Above 40', 'Actual Below 40')
results = df %>% select(Time, predict, NO2, predictOver40, actualOver40)
table(results$predictOver40, results$actualOver40)


# accurately predict above 40 when actual is above 40   ***** GOOD NEWS IF THE MODEL IS GOOD AT THIS  ---- NEED TO MAXIMIZE *****
accurate_pred_over_40 = table(results$predictOver40, results$actualOver40)[1,1] / (table(results$predictOver40, results$actualOver40)[1,1] + table(results$predictOver40, results$actualOver40)[2,1])
paste0('Accurately predict NO2 above 40: ',round(accurate_pred_over_40,3))

# accurate predict below 40 when actual is below 40
accurate_pred_below_40 = table(results$predictOver40, results$actualOver40)[2,2] / (table(results$predictOver40, results$actualOver40)[1,2] + table(results$predictOver40, results$actualOver40)[2,2])
paste0('Accurately predict NO2 below 40: ',round(accurate_pred_below_40,3))

# inaccurately predict above 40 when actual is below 40
inaccurate_pred_above_40 = table(results$predictOver40, results$actualOver40)[1,2] / (table(results$predictOver40, results$actualOver40)[1,1] + table(results$predictOver40, results$actualOver40)[1,2])
paste0('Inaccurately predict NO2 above 40: ',round(inaccurate_pred_above_40,3))

# inaccurate predict below 40 when actual is above 40 ****BIG PROBLEM IS THIS IS THE CASE ---- NEED TO MINIMIZE THIS PERCENTAGE ******
inaccurate_pred_below_40 = table(results$predictOver40, results$actualOver40)[2,1] / (table(results$predictOver40, results$actualOver40)[2,1] + table(results$predictOver40, results$actualOver40)[1,1])
paste0('Inaccurately predict NO2 below 40: ',round(inaccurate_pred_below_40,3))

```



# VALI SET - (with reset at 1pm)
# 1. LAGS 1-3, CURRENT METEO, SLOPES PRIOR 8AM
# MAE & RESULTS TABLE FOR VALIDATION
```{r}
# REPEAT FOR VALIDATION DATA SET
infilePath = file.path("~/Dropbox","NU","THESIS","DATASETS","RESULTS", "PREDICTIONS","ONE_HOUR_AHEAD_PREDICTIONS")
df         = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_1_hr_ahead.csv"),na.strings=c("NA"," "))
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_2_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_3_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_4_hr_ahead.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_5_hr_ahead_reset_13.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_6_hr_ahead_reset_13.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_7_hr_ahead_reset_13.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
dfNew      = read.csv(file.path(infilePath,"greenwich_eltham_NO2_ann_vali_8_hr_ahead_reset_13.csv"),na.strings=c("NA"," "))
df = rbind(df, dfNew)
df

# CALC the MAE here
# VALIDATION: evaluate the quality of the prediction using the test data set
df.vali.mae = mean(abs(df$predict-df$NO2))
df.vali.mae          # 5.847267 NO2


# CONFUSION MATRIX
df$predictOver40 = ifelse(df$predict > 40, 'Predict Above 40', 'Predict Below 40')
df$actualOver40 = ifelse(df$NO2 > 40,  'Actual Above 40', 'Actual Below 40')
results = df %>% select(Time, predict, NO2, predictOver40, actualOver40)
table(results$predictOver40, results$actualOver40)


# accurately predict above 40 when actual is above 40   ***** GOOD NEWS IF THE MODEL IS GOOD AT THIS  ---- NEED TO MAXIMIZE *****
accurate_pred_over_40 = table(results$predictOver40, results$actualOver40)[1,1] / (table(results$predictOver40, results$actualOver40)[1,1] + table(results$predictOver40, results$actualOver40)[2,1])
paste0('Accurately predict NO2 above 40: ',round(accurate_pred_over_40,3))

# accurate predict below 40 when actual is below 40
accurate_pred_below_40 = table(results$predictOver40, results$actualOver40)[2,2] / (table(results$predictOver40, results$actualOver40)[1,2] + table(results$predictOver40, results$actualOver40)[2,2])
paste0('Accurately predict NO2 below 40: ',round(accurate_pred_below_40,3))

# inaccurately predict above 40 when actual is below 40
inaccurate_pred_above_40 = table(results$predictOver40, results$actualOver40)[1,2] / (table(results$predictOver40, results$actualOver40)[1,1] + table(results$predictOver40, results$actualOver40)[1,2])
paste0('Inaccurately predict NO2 above 40: ',round(inaccurate_pred_above_40,3))

# inaccurate predict below 40 when actual is above 40 ****BIG PROBLEM IS THIS IS THE CASE ---- NEED TO MINIMIZE THIS PERCENTAGE ******
inaccurate_pred_below_40 = table(results$predictOver40, results$actualOver40)[2,1] / (table(results$predictOver40, results$actualOver40)[2,1] + table(results$predictOver40, results$actualOver40)[1,1])
paste0('Inaccurately predict NO2 below 40: ',round(inaccurate_pred_below_40,3))

```


# DRAW PLOTLY OF DAILY PREDICT VS ACTUAL FOR DAYS WHEN OVER 40 ... 
# DRAW TABLE WITH NUMBER OF TIMES >40 WHERE PREDICT ACCURATE
```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[1])

good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
good.predict$text = 'Accurate Prediction'
#bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
#bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[1])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```




```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[2])

good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
good.predict$text = 'Accurate Prediction'
bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[2])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```



```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[3])

good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
good.predict$text = 'Accurate Prediction'
#bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
#bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[3])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```




```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[4])

#good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
#good.predict$text = 'Accurate Prediction'
bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[4])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```



```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[5])

good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
good.predict$text = 'Accurate Prediction'
bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[5])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```



```{r}
library(plotly)
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Hour of Day",
  titlefont = f
)
y <- list(
  title = "Average Hourly NO2",
  titlefont = f
)


# get days where >1 hour has over 40 NO2.
results$Date   = as.Date(results$Time)
Over40AllObs   = results  %>% filter(actualOver40 == 'Actual Above 40')
DfOver40AllDates = data.frame(unique(Over40AllObs$Date))
colnames(DfOver40AllDates) = c('Date')

resultsOver40AllDateObs      = inner_join(results, DfOver40AllDates, by=c("Date" = "Date"))
resultsOver40AllDateObs$Time = substr(resultsOver40AllDateObs$Time, 1, 16)
resultsOver40AllDateObs$Time = as.POSIXct(resultsOver40AllDateObs$Time)
resultsOver40AllDateObs[ order(resultsOver40AllDateObs$Time , decreasing = FALSE ),]  # sort data frame in appropriate order

# date 1: annotate graph when correctly or incorrectly predict over 40
data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[6])

#good.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Above 40')
#good.predict$text = 'Accurate Prediction'
bad.predict = data %>% filter(actualOver40 == 'Actual Above 40' & predictOver40=='Predict Below 40')
bad.predict$text = 'Inaccurate Prediction'
euLimit = rep(40,8)

m = rbind(good.predict, bad.predict)

a <- list(
  x = m$Time,
  y = m$NO2,
  text = m$text,   #'fail predict',
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 6,
  ax=1
)

data = resultsOver40AllDateObs %>% filter(resultsOver40AllDateObs$Date == unique(Over40AllObs$Date)[6])
p <- plot_ly(data, x = ~Time, y=~NO2, name = 'Actual NO2', type='scatter', mode='lines+markers') %>%      
  add_trace(y = ~predict, name = 'Predicted NO2',mode = 'lines+markers') %>%
  add_trace(y = ~euLimit, name = 'EU Limit', line = list(color = 'rgb(205, 12, 24)', width = 4, dash = 'dash')) %>%
  layout(annotations=a, title = paste0("Actual vs Predicted Hourly NO2 Greenwich Eltham: ",unique(Over40AllObs$Date)[1]), 
  xaxis = list(title  = "Hour of Day"),
  yaxis = list (title = "Hourly NO2 Concentration (µg/m3)"))
p
```












