---
title: "thesis random forest"
author: "Steven Futter"
date: "5/7/2017"
output: html_document
---

---
title: "time-series forecasting"
author: "Steven Futter"
date: "5/6/2017"
output: html_document
---


NOtes: 
given that the ex post multiple linear regression variable selection methods yield small MAE's and the ex ante forecast quite large we have confirmed that the challenge here is to ensure that we can accurately forecast the future values for each predictor so that the multiple linear regression can be used to predict the response at a future hour of the school day.

Let's look at using an ARIMA model to predict the predictor variables at intervals throughout the day, then use those values for inputs into the multiple linear regression models already derived in the thesis_data_model.Rmd tab. 


https://stats.stackexchange.com/questions/125909/forecast-using-arima-models

```{r}
library(dplyr)
inPath = file.path("~/Dropbox","NU","THESIS","DATASETS","GROOMED") # Home path
df = read.csv(file.path(inPath,"greenwich_eltham_imputed_school_hours.csv"),na.strings=c("NA"," "))

codePath = file.path("~/Dropbox","NU","THESIS","R CODE")
source(file.path(codePath,"thesis_data_preparation_part3.R"))

df$Date = as.Date(df$Date)

df.train = subset(df, Date <= '2013-05-28')
df.test  = subset(df, Date > '2013-05-28')

head(df.train)
tail(df.train)

head(df.test)
tail(df.test)

dim(df.train)
dim(df.test)

df.train = df.train %>% dplyr::select(Date:lag.3.WSPD)
df.test  = df.test  %>% dplyr::select(Date:lag.3.WSPD)


```

# START RANDOM FOREST HERE
https://cran.r-project.org/web/packages/randomForest/randomForest.pdf

Use the random forest to predict the future levels of school hour pollutants and meteorological variables.

Goal: need to be able to predict the level of pollution at different intervals throughout the day...


```{r}
library(randomForest)
set.seed(1)

# get columns to which can apply ARIMA model first
df1 = df %>% dplyr::select(NO:lag.3.WSPD)

# Function that returns Mean Absolute Error
mae <- function(error)
{
    mean(abs(error))
}


mse <- function(error)
{
  mean(error^2)
}


# for (i in 1:length(names(df1))) {
#for (i in 1:length(names(df1))){

response = 'NO2' #names(df1)[i] #'NO'

df.train 

randomForest.fit = randomForest(response~ ., data=df.train, mtry=5, importance=TRUE)
randomForest.fit

yhat.randomForest.train = predict(randomForest.fit, newdata = df.train)
randomForest.mse.train = mean((yhat.randomForest.train-df.train$NO2)^2)  # MSE = 0.006340628 (in-sample)
randomForest.mse.train

yhat.randomForest.test = predict(randomForest.fit, newdata = df.test)
randomForest.mse.test = mean((yhat.randomForest.test-df.test$NO2)^2)         # MSE = 0.0457037 (out-of-sample)
randomForest.mse.test





# Fit model to train data
fit  <- Arima(df.train[,c(response)], order = c(arimaorder(arima.model)[1],arimaorder(arima.model)[2],arimaorder(arima.model)[3])) 

# Apply fitted model to the test data
fit2 <- Arima(df.test[,c(response)], model = fit)        

arima.fitted.test.vals = fitted(fit2)

df.test = data.frame(df.test, arima.fitted.test.vals)
names(df.test)[names(df.test)=="arima.fitted.test.vals"] <- paste0('arima.fitted.',response)

}

head(df.test)

outfilePath = file.path("~/Dropbox","NU","THESIS","DATASETS","RESULTS")
write.csv(df.test, file.path(outfilePath,'results_test_data_all_vars_arima.csv'))

# #################################
# # MAE and MSE and RMSE etc... ###
# #################################
# error = arima.fitted.test.vals - df.test[,c(response)]
# results = mae(error) 
# 
# infilePath  = file.path("~/Dropbox","NU","THESIS","DATASETS","RESULTS")
# outfilePath = file.path("~/Dropbox","NU","THESIS","DATASETS","RESULTS")
# 
# ############################################################
# ########################## CAREFUL #########################
# ############################################################
# # ONLY RUN FIRST TIME TO RESET RESULTS TABLE 
# # head(results)
# #write.csv(results, file.path(outfilePath,'results_arima.csv'))
# # 
# # # second time through - append to end #
# # head(results)
# resultsImport = read.csv(file.path(infilePath,"results_arima.csv"),na.strings=c("NA"," "))
# resultsImport = resultsImport[,2:length(names(resultsImport))]  # remove the 'X' column
# resultsAppended = rbind(resultsImport, results)
# write.csv(resultsAppended, file.path(outfilePath,'results_arima.csv'))
#
#}
```


```{r}




# # STL Decomposition: seasonal decomposition of time series by LOESS
# # Break down the time series into trend, season, and irregular (noise)
# ts = ts(df$NO2, start=2010-09-27, frequency=11)
# fit = stl(ts, s.window='periodic')
# plot(fit)
# df

auto.arima(df$NO2) # ARIMA(3,1,4)

# Fit model to train data
fit <- Arima(df.train$NO2, order = c(3,1,4)) # , xreg = x)

# Fit same model to test data
fit2 <- Arima(y2, model = fit, xreg = x2)


# fit.rhum = auto.arima(df$RHUM)
# summary(fit.rhum)
# predict(fit.rhum,n.ahead = 2)


# Plot fitted values
plot(fitted(fit))
lines(fitted(fit2), col = "blue")

   




# ## without drift
# fit.m <- arima(df$NO2, order = c(0,1,1))
# forecast.m <- plot(forecast(fit.m, h=10))
# # or
# predict(fit.m, n.ahead = 8)

```


```{r}
library(randomForest)
set.seed(1)

randomForest.fit = randomForest(NO2~ ., data=df.train, mtry=5, importance=TRUE)
randomForest.fit

yhat.randomForest.train = predict(randomForest.fit, newdata = df.train)
randomForest.mse.train = mean((yhat.randomForest.train-df.train$NO2)^2)  # MSE = 0.006340628 (in-sample)
randomForest.mse.train

yhat.randomForest.test = predict(randomForest.fit, newdata = df.test)
randomForest.mse.test = mean((yhat.randomForest.test-df.test$NO2)^2)         # MSE = 0.0457037 (out-of-sample)
randomForest.mse.test


predictions=yhat.randomForest.train
error <- predictions-df.train$NO2
rmse.train  = rmse(error)   #eltham 4.930543
mse.train   = mse(error)    #heathrow 203.9976
mae.train   = mae(error)    #heathrow 11.54623
r2.train    = r2(df.train$NO2,predictions)   #actual, predicted      #heathrow 0.5780822
rmsle.train = msaenet.rmsle(df.train$NO2,predictions)   #heathrow 0.6629021

model                   = c('Random Forest')
type                    = c('Training')
results.train           = data.frame(model,type,rmse.train, mse.train,mae.train,r2.train,rmsle.train)
colnames(results.train) = c('Model','Data Set','RMSE','MSE','MAE','R2','RMSLE')
results.train

# Calculate testing error
predictions.test = yhat.randomForest.test
test.error <- predictions.test-df.test$NO2
rmse.test   = rmse(test.error)   #heathrow 14.57655
mse.test    = mse(test.error)    #heathrow 212.4759
mae.test    = mae(test.error)    #heathrow 11.71268
r2.test     = r2(df.test$NO2,predictions.test) #actual, predicted    #heathrow 0.5161738
rmsle.test  = msaenet.rmsle(df.test$NO2,predictions.test)   #heathrow 0.696822

type                   = c('Testing')
results.test           = data.frame(model,type,rmse.test, mse.test,mae.test,r2.test,rmsle.test)
colnames(results.test) = c('Model','Data Set','RMSE','MSE','MAE','R2','RMSLE')
results.test

results = rbind(results.train,results.test)
results


```